### 2023-11-13 Manuscript/manuscript.tex
Initial commit
Notes are not updated.

\documentclass[a4paper]{siamart190516}

\usepackage{damacros}

% Sets running headers as well as PDF title and authors
\headers{Preliminary notes}{Daniele, Sage, and Zack}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.
\title{Notes on RBFs methods for Neural Fields}

% Authors: full names plus addresses.
\author{%
  Daniele 
  \and
  Sage 
  \and 
  Zack 
  % \thanks{%
  %   Vrije Universiteit Amsterdam,
  %   Department of Mathematics,
  %   Faculteit der Exacte Wetenschappen,
  %   De Boelelaan 1081a,
  %   1081 HV Amsterdam, The Netherlands.
  % \protect\\
  %   Inria Sophia Antipolis M\'editerran\'ee Research Centre,
  %   MathNeuro Team,
  %   2004 route des Lucioles-Boîte Postale 93 06902,
  %   Sophia Antipolis, Cedex, France.
  % \protect\\
  %   (\email{d.avitabile@vu.nl}, \url{writemywebpage}).
  % }
  % % \and
% %   Paul T. Frank \thanks{Department of Applied Mathematics, Fictional University, Boise, ID 
% % (\email{ptfrank@fictional.edu}, \email{jesmith@fictional.edu}).}
% % \and Jane E. Smith\footnotemark[3]
}

\graphicspath{{./Figures/}}

\begin{document}

\maketitle

\begin{abstract}
  Preliminary notes on RBFs for neural fields
\end{abstract}

\section{Example of collocation scheme for a neural field on a ring} 
Let us consider a neural field model on a ring, with cosine synaptic kernel, 
\begin{equation}\label{eq:NFRing}
  \begin{aligned}
 & \partial_t u(x,t) =  -u(x,t) + g(x,t)+ \int_{-\pi}^{\pi} \cos(|x-x'|) f(u(x',t)) dx',
   \qquad x \in [-\pi,\pi) \\
 & u(x,0) = v(x) 
  \end{aligned}
\end{equation}
where $g$, $f$, and $v$ are prescribed functions, and let us derive an intuitive scheme
for this problem. We pick $n+1$ evenly spaced points in $[-\pi,\pi]$
\begin{equation}\label{eq:nodesRing}
  x_i = -\pi + h i, \qquad h = \frac{2\pi}{n}, \qquad i = 0, \ldots, n.
\end{equation}
Because of periodicity, we expect to need only $n$ of those $n+1$ points, and indeed
this turns out to be the case. If $v(x)$ is a periodic function, applying the
trapezium rule with the grid above gives (note the sum runs over just $n$ points)
\[
  \int_{-\pi}^{\pi} v(x')\,d x' \approx \sum_{j=1}^{n} v(x_j) h.
\]
The integrand in \cref{eq:NFRing} is periodic in $x'$, so we evaluate the neural
field at $x_1, \ldots, x_n$ ($x_0$ excluded) and simultaneously apply the quadrature rule
above to obtain
\[
  \begin{aligned}
  & U'_i(t) = -U_i(t) + \sum_{j=1}^{n} \cos(x_i - x_j) f(U_j(t)) h + G_i(t), \qquad i = 1,\ldots,n  \\
  & U_i(0) = V_i
  \end{aligned}
\]
that is, a system in the form 
\begin{equation}\label{eq:NFODE}
  \begin{aligned}
   &  U'(t) = - U(t) + M F(U(t)) +  G(t)\\ 
   &  U(0) = V
  \end{aligned}
\end{equation}  
in which $U(t), G(t), F(U(t)), V$ are vectors in $\RSet^n$ for all $t$, and in which
$M$ is the $n$-by-$n$, full matrix entries $M_{ij} = \cos(x_i - x_j)h$. The intention
here is to use $U_i(t)$ as an approximation for $u(x_i,t)$.

The scheme above is an example of \textit{collocation scheme}. Even though this may
not be apparent from the quick presentation above, it scaffolding is polynomial
approximation, and more specifically polynomial interpolation. If we
look at the scheme with the right lenses, we find that many important properties of
the scheme revolve around solving following problem: 
\begin{problem}[Interpolation problem]\label{prob:interpolation}
Given a function $v(x)$, and a set of $n$ points $\{ x_i \}_{i=1}^n$, find a function
$v_n(x)$ such that 
\[
  v(x_i) = v_n(x_i), \qquad i = 1,\ldots, n.
\]  
Once $v_n$ is constructed, study how well $v_n(x)$ approximates $v(x)$ away from the
nodes $x_i$, for instance by studying $\| v - v_n \|$ as $n \to \infty$ in a suitable
norm.
\end{problem}
The problem above is cast in terms of a generic function $v(x)$, but thinking of this
function $v$ as being the initial condition of the neural field \cref{eq:NF} is
useful.

In \cite{avitabileProjectionMethodsNeural2023} it is shown that the scheme discussed
above stems from the following choices: 
\begin{enumerate}
  \item The points are evenly spaced in $[-\pi,\pi]$.
  \item The function $v_n$ that subtends the scheme is of the following form:
    \begin{equation}\label{eq:LagrangeInterp}
      v_n(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x), 
    \end{equation}
    where $\ell_i(x)$ are Lagrange polynomials, hence $\ell_i(x_j) = \delta_{ij}$.
    This choice of $v_n$ is one way to solve the interpolation
    \cref{prob:interpolation}.
\end{enumerate}
In \cite{avitabileProjectionMethodsNeural2023} it is also shown that a variety of
other schemes, of Finite Elements or Galerkin type, are analysable following a common
framework.

\subsection{RBFs for Neural field equations}\label{ssec:RBFMotivation} From a
numerical standpoint, the discussion above brings about the following questions:
  \begin{itemize}
    \item How to choose a parsimonious distributions of points? Under which conditions
      is a homogeneous point distribution a good idea? What if we are in multiple
      spatial dimensions? What if the cortex has curvature, as a manifold in 3D? How
      to handle refinements?
    \item The interpolation function $v_n$ given in \cref{eq:LagrangeInterp} is one
      of many possible choices. Can we gain by changing the form of $v_n$, and
      perhaps pairing it to a point distribution that ``goes well with $v_n$"?
    \item Evaluating the right-hand side of system \cref{eq:NFODE} involves a
      matrix-vector multiplication with a dense matrix $M$. This means the scheme
      won't scale well on large cortices, because at each time step it takes
      $O(n^2)$ operations to evaluate the right-hand side. Can we choose points $\{
      x_i \}$, interpolant $v_n$, and quadrature rules so that $M$ is sparse?
  \end{itemize}

  Using RBFs is promising and interesting because of the following points:
  \begin{itemize}
      \item RBF handle point distributions in a flexible and adaptable manner. They
        work well in multiple spatial dimensions, and on curved geometries. It is is
        relatively simple to add points where it is needed; RBFs are constructed
        using translates of a kernel function, without a mesh.
      \item Convergence results for RBFs seem available in literature, so there is
        hope to connect to the framework in
        \cite{avitabileProjectionMethodsNeural2023} and quickly obtain convergence
        rates.
      \item We will focus on RBF interpolants of the form
        \begin{equation}\label{eq:RBFInterpolation}
          v_n(x) = \sum_{i=1}^{n} \sum_{j=1}^{k_i} V_{ij} \Phi(\| x - \xi_{ij} \|),
        \end{equation}
        where $\xi_{ij}$ are suitably chosen subsets of $\{ x_i \}$, as we will
        specify later. This form of interpolation induces a \textit{sparsification}
        of the matrix $M$.
  \end{itemize}

\subsection{Main objectives of this work}\label{ssec:Objectives} An interesting
self-contained publication could be centred around the following:
\begin{itemize}
  \item Describe collocation schemes for generic neural fields posed on surfaces,
    using RBFs as the underlying interpolants.
  \item If at all possible, infer the convergence rate of these schemes. If not,
    present an empirical convergence scheme on a simple model such as
    \cref{eq:NFRing} (or perhaps a 2D version of it).
  \item Showcase the method on a realistic cortical geometry, using either a
    multi-layer neural field, or a Montbrio version of neural field.
  \item If we want, give people simple routines to solve neural fields.
\end{itemize}
Possibly in a separate publication, we could look at the problem with diffusion, and
time adaptivity.

\section{Heuristic scheme formulation}\label{sec:model} To understand how RBFs enter
into the discretisation, we show how interpolation determines the collocation scheme,
that is, how the choice of $v_n$ translates into a collocation scheme. The purpose of
this section is to prepare us for that: no mention will be made yet about
interpolation, or $v_n$, but we are getting ready for that.

Let us consider a neural field posed on
a compact domain $D \subset \RSet^d$
\begin{equation}\label{eq:NF}
  \begin{aligned}
  &
  \begin{aligned}
    \partial_t u(x,t) =  -u(x,t)  &+ g(x,t) \\
                & + \int_{D} w(x,x') f(u(x',t))d \mu(x'),
  \end{aligned}
  && (x,t) \in D \times [0,T] \\
  & u(x,0) = v(x) && x \in D 
  \end{aligned}
\end{equation}
It is assumed here that the integral is on a manifold, hence it is expressed in terms
of a measure $\mu$. Heuristically, a collocation scheme for \cref{eq:NF} can be conceived
in steps, as follows:

\textbf{Step 1: nodes selection.} Pick $n$ points in the domain $D$. We denote them by
$\Xi = \{ x_i \}_{i \in \NSet_n} \subset D$. \da[]{I have used $\NSet_n = \{ 1,\ldots,n
\}$. I also dislike $\Xi$, we can change it later} In what follows, it may be useful to make $n$ depend on a positive real number
$h$, a discretisation parameter, so that $n(h) \to \infty$ as $h \to 0$, or as $h \to
\infty$. An example of such relationship is given in \cref{eq:nodesRing}. We will
omit the dependence of $n$ on $h$ when possible.

\textbf{Step 2: Collocation.} 
Impose that \cref{eq:NF} holds at each node (\textit{collocate the equation} at
each $x_i \in \Xi$). This means replacing \cref{eq:NF} by a set of $n$ equations
\begin{equation}\label{eq:NFColl}
  \begin{aligned}
  &
  \begin{aligned}
    \partial_t u_n(x_i,t) =  -u_n(x_i,t) &+ g(x_i,t) \\
                & + \int_{D} w(x_i,x') f(u_n(x',t))d \mu(x'),
  \end{aligned}
  && (x_i,t) \in \Xi \times [0,T] \\
  & u_n(x_i,0) = v(x_i) && x_i \in \Xi 
  \end{aligned}
\end{equation}

Comparing \cref{eq:NF} with \cref{eq:NFColl} we note that the latter is expressed in
terms of the unknown $u_n$, which is an approximation to $u$: the two functions
satisfy evolution equations that coincide only at the nodes $\Xi$, not on the whole $D$.
Another observation is that \cref{eq:NFColl} may be useful for theoretical
estimates, but it is clearly not implementable on a computer as is, because integrals must be
approximated. This takes us to the last step.

\textbf{Step 3: Quadrature.} Select a quadrature for the integral, and solve
\begin{equation}%\label{eq:NFODE}
  \begin{aligned}
   & \tilde U'(t) = -\tilde U(t) + K(\tilde U(t)) + \tilde G(t) \qquad  t \in [0,T]\\ 
   & \tilde U(0) = V
  \end{aligned}
\end{equation}  
where $\tilde U(t), G(t), K(\tilde U(t)) \in \RSet^n$ for all $t$, and $\tilde U_i(t)$ approximates
$u_n(x_i,t)$. The nonlinear mapping $K(U)$ involves a matrix-vector multiplication,
and incorporates quadrature weights.

\section{Collocation schemes as interpolatory projection schemes}\label{sec:projectionSchemes} 
\da[inline]{This section can be skipped at a first pass}

Once we have refined the steps of a collocation scheme, we can bring interpolation
into the picture. In \cite{avitabileProjectionMethodsNeural2023} it is shown that convergence estimates
for the collocation scheme (and other schemes) are possible disregarding step 3. In
fact, with just steps 1-2 can estimate how fast $u_n$ tends to $u$, as $n \to \infty$.
Once a convergence rate is established, then this instructs which quadrature scheme
to select: the quadrature should respects this rate, so that we can use $\tilde U_i$ as an
approximation to $u(x_i,t)$, not simply $u_n(x_i,t)$.

We do not have to get into the details of
\cite{avitabileProjectionMethodsNeural2023}, but one aspect is important for these
notes: it is useful to understand how \cref{eq:NFColl} is derived from \cref{eq:NF}
using a projection. The RBFs will provide, in fact, a flexible and accurate
\textit{interpolatory projector}.

To make progress on this, let us assume we seek for a solution to the neural field
equations in the space of continuous functions $\XSet = C(D)$. A projector $P_n$ is a
linear mapping from $\XSet$ to an $n$-dimensional subspace $\XSet_n \subset \XSet$
such that $P_n (P_n v) = P_n v$ for all $v \in \XSet$. The distinctive feature of an
\textit{interpolatory projector} is that it has the property  
\[
  (P_n v)(x_i) = v(x_i), \qquad x_i \in \Xi,
\] 
that is, $P_n v$ is a function of $x$, which approximates $v(x)$, and it is exact
at the nodes $x_i$. The previous sections contain two interpolatory projectors in
disguise. In \cref{eq:LagrangeInterp} we wrote
\[
  v_n(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x). 
\]
Now $v$ is a continuous function so $v \in \XSet$, and $v_n$ is a linear combination
of $\{ \ell_j \}_{j=1}^n$, so $v_n \in \XSet_n = \spn \{ \ell_1, \ldots, \ell_n \}$.
This means we can interpret the previous writing as
\[
  (P_n v)(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x),
\]
that is, $v_n$ is the action of $P_n$ on the function $v$, $v_n = P_n v$.
The same can be said about \cref{eq:RBFInterpolation}, which defines another choice
of the interpolant. Another choice of projector $P_n$.

Writing things in terms of $P_n$ is useful, as it compactifies the notation. Firstly,
it is possible to rewrite the neural field equation \cref{eq:NF} as the following
Cauchy problem on the Banach space $\XSet$
\begin{equation}\label{eq:cauchy}
  \begin{aligned}
  & \dot U(t) = - U(t) + WF(U(t)sketch.pnsketch.pngg) + G(t) =:  N(U(t),t), \qquad t \in J, \\
  & U(0) = V.
  \end{aligned}
\end{equation}
The identification with \cref{eq:NF} is established with the positions
\[
  \begin{aligned}
   & V = v(\blank)\\
   & U \colon [0,T] \to \XSet, && t \mapsto u(\blank,t) \\
   & G \colon [0,T] \to \XSet, && t \mapsto g(\blank,t) \\
   & W \colon \XSet \to \XSet, && v \mapsto \int_{D} w(\blank,y) v(y) \,d \mu(x') \\
   & F \colon \XSet \to \XSet, && v \mapsto f(v). 
  \end{aligned}
\]
The Cauchy problem is well-posed under general assumptions that is, there is a unique
$u \in C^1(J,\XSet)$ satisfying \cref{eq:cauchy}.

Secondly, it is possible to interpret \cref{eq:NFColl} in operator form, as
\begin{equation}\label{eq:cauchyPn}
  \begin{aligned}
  & \dot U_n(t) = - U_n(t) + P_n WF(U(t)) + P_n G(t) =  P_n N(U_n(t),t), \qquad t \in J, \\
  & U_n(0) = P_n V,
  \end{aligned}
\end{equation}
which is a Cauchy problem on the space $\XSet_{n}$.

\section{RBF scheme for neural field equations}\label{sec:RBFScheme} 
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{sketch}
  \caption{The patch $D_i$ of the spatial domain $D$ contains $k_i =4$}
  \label{fig:sketch}
\end{figure}

\section{Conventions} \label{sec:conventions}
See \texttt{docsiamart.pdf} for many examples. I generally adopt the conventions of
that template. I remark a few things below.

\subsection{References} \label{sec:references}
These are random citations \cite{KoMa14,siam}. For cross-references I use
\texttt{cleveref}: like this
\cref{eq:test}, like this \Cref{eq:test}, or like this \cref{sec:conventions}, or
this \cref{hyp:trivial}, or this \cref{clm:trivial}.
\begin{equation}\label{eq:test}
  x = y
\end{equation}

\subsection{Todos}
You can find templates of other author's todos in
\texttt{damacros.sty}, which you are welcome to edit.
This is how I make a short comment\da{Short comment by Daniele}. Comments can also go
inline like the one below.
\da[inline]{Long comments by Daniele} 
% \missingfigure
% 
% \section*{Acknowledgments}

\section{Testing commands}

\subsection{Testing maths operators}
Here are tests for operators
\[
  \spec, \quad 
  \range, \quad 
  \spn, \quad
  \diag, \quad
  \meas, \quad
  \id, \quad
  \real z, \quad
  \imag z, \quad
  \e, \quad
  \iunit, \quad
\]

\subsection{Testing additional theorem-like environments} The following additional
theorem-like environments may be useful.

\begin{hypothesis} \label{hyp:trivial}
  This is a hypothesis
\end{hypothesis}

\begin{claim} \label{clm:trivial}
  This is a claim
\end{claim}

\subsection{Testing sets}
This are a few examples of sets (see more commands in \texttt{damacros.sty})
\[
  \RSet, \quad
  \CSet, \quad
  \ZSet, \quad
  \sph, \quad
  \ldots
\]

\subsection{Empty arguments}
Command for empty arguments, which require spacing
\[
  u(\blank,t) \quad t \in \RSet
\]

\subsection{Calligraphic fonts}
If one really must use calligraphic fonts, there are macros for this.
\[
  \calA, \quad
  \calL, \quad
  \calM, \quad
  \calS, \quad
\]

\bibliographystyle{siamplain}
\bibliography{references}
\end{document}

### 2023-11-14 Manuscript/manuscript.tex
Added section on projection methods

@@
   \begin{aligned}
- & \partial_t u(x,t) =  -u(x,t) + g(x,t)+ \int_{-\pi}^{\pi} \cos(|x-x'|) f(u(x',t)) dx',
+ & \partial_t u(x,t) =  -u(x,t) + g(x,t)+ \int_{-\pi}^{\pi} \cos(|x-y|) f(u(y,t)) dy,
    \qquad x \in [-\pi,\pi) \\
@@
 \[
-  \int_{-\pi}^{\pi} v(x')\,d x' \approx \sum_{j=1}^{n} v(x_j) h.
+  \int_{-\pi}^{\pi} v(y)\,d y \approx \sum_{j=1}^{n} v(x_j) h.
 \]
-The integrand in \cref{eq:NFRing} is periodic in $x'$, so we evaluate the neural
+The integrand in \cref{eq:NFRing} is periodic in $y$, so we evaluate the neural
 field at $x_1, \ldots, x_n$ ($x_0$ excluded) and simultaneously apply the quadrature rule
@@
 that is, a system in the form 
-\begin{equation}\label{eq:NFODE}
+\begin{equation}
   \begin{aligned}
@@
 
-In \cite{avitabileProjectionMethodsNeural2023} it is shown that the scheme discussed
+In \cite{Avitabile:2023ab} it is shown that the scheme discussed
 above stems from the following choices: 
@@
 \end{enumerate}
-In \cite{avitabileProjectionMethodsNeural2023} it is also shown that a variety of
+In \cite{Avitabile:2023ab} it is also shown that a variety of
 other schemes, of Finite Elements or Galerkin type, are analysable following a common
@@
         hope to connect to the framework in
-        \cite{avitabileProjectionMethodsNeural2023} and quickly obtain convergence
+        \cite {Avitabile:2023ab} and quickly obtain convergence
         rates.
@@
 
+\section{Notation}\label{sec:notation} We will use the index sets by $\NSet_n = \{
+1,\ldots, n\}$. In addition, we will endow $\RSet^n$ with the standard Euclidean
+norm, which we shall denote by $\|\blank\|_2$, or simply $\| \blank \|$.
+
 \section{Heuristic scheme formulation}\label{sec:model} To understand how RBFs enter
@@
     \partial_t u(x,t) =  -u(x,t)  &+ g(x,t) \\
-                & + \int_{D} w(x,x') f(u(x',t))d \mu(x'),
+                & + \int_{D} w(x,y) f(u(y,t))d \mu(y),
   \end{aligned}
@@
 \textbf{Step 1: nodes selection.} Pick $n$ points in the domain $D$. We denote them by
-$\Xi = \{ x_i \}_{i \in \NSet_n} \subset D$. \da[]{I have used $\NSet_n = \{ 1,\ldots,n
-\}$. I also dislike $\Xi$, we can change it later} In what follows, it may be useful to make $n$ depend on a positive real number
-$h$, a discretisation parameter, so that $n(h) \to \infty$ as $h \to 0$, or as $h \to
-\infty$. An example of such relationship is given in \cref{eq:nodesRing}. We will
-omit the dependence of $n$ on $h$ when possible.
+$\Xi = \{ x_i \}_{i \in \NSet_n} \subset D$. In what follows, it may be useful to
+make $n$ depend on a positive real number $h$, a discretisation parameter, so that
+$n(h) \to \infty$ as $h \to 0$, or as $h \to \infty$. An example of such relationship
+is given in \cref{eq:nodesRing}. We will omit the dependence of $n$ on $h$ when
+possible.
 
@@
     \partial_t u_n(x_i,t) =  -u_n(x_i,t) &+ g(x_i,t) \\
-                & + \int_{D} w(x_i,x') f(u_n(x',t))d \mu(x'),
+                & + \int_{D} w(x_i,y) f(u_n(y,t))d \mu(y),
   \end{aligned}
@@
 \textbf{Step 3: Quadrature.} Select a quadrature for the integral, and solve
-\begin{equation}%\label{eq:NFODE}
+\begin{equation}\label{eq:NFODEQuad}
   \begin{aligned}
-   & \tilde U'(t) = -\tilde U(t) + K(\tilde U(t)) + \tilde G(t) \qquad  t \in [0,T]\\ 
-   & \tilde U(0) = V
+   & \tilde \alpha'(t) = -\tilde \alpha(t) + \tilde K(\tilde \alpha(t)) + \gamma(t) \qquad  t \in [0,T],\\ 
+   & \tilde \alpha (0) = \nu,
   \end{aligned}
 \end{equation}  
-where $\tilde U(t), G(t), K(\tilde U(t)) \in \RSet^n$ for all $t$, and $\tilde U_i(t)$ approximates
-$u_n(x_i,t)$. The nonlinear mapping $K(U)$ involves a matrix-vector multiplication,
-and incorporates quadrature weights.
+where $\tilde \alpha(t)$, $\gamma(t), \tilde K(\tilde \alpha(t)), \nu \in \RSet^n$ with
+$\gamma_i(t) = g(x_i,t)$, $\nu_i = v(x_i)$, and where $(\tilde \alpha(t))_i$
+approximates $u_n(x_i,t)$. The nonlinear mapping $K(\tilde \alpha)$ involves a matrix-vector
+multiplication, and incorporates quadrature weights.
 
 \section{Collocation schemes as interpolatory projection schemes}\label{sec:projectionSchemes}
-\da[inline]{This section can be skipped at a first pass}
 
 Once we have refined the steps of a collocation scheme, we can bring interpolation
-into the picture. In \cite{avitabileProjectionMethodsNeural2023} it is shown that convergence estimates
+into the picture. In \cite{Avitabile:2023ab} it is shown that convergence estimates
 for the collocation scheme (and other schemes) are possible disregarding step 3. In
@@
 Once a convergence rate is established, then this instructs which quadrature scheme
-to select: the quadrature should respects this rate, so that we can use $\tilde U_i$ as an
-approximation to $u(x_i,t)$, not simply $u_n(x_i,t)$.
+to select: the quadrature should respects this rate, so that we can use $(\tilde
+\alpha(t))_i$ as an approximation to $u(x_i,t)$, not simply $u_n(x_i,t)$.
 
 We do not have to get into the details of
-\cite{avitabileProjectionMethodsNeural2023}, but one aspect is important for these
+\cite{Avitabile:2023ab}, but one aspect is important for these
 notes: it is useful to understand how \cref{eq:NFColl} is derived from \cref{eq:NF}
@@
 \]
-Now $v$ is a continuous function so $v \in \XSet$, and $v_n$ is a linear combination
+Since $v$ is a continuous function, then so $v \in \XSet$, and $v_n$ is a linear combination
 of $\{ \ell_j \}_{j=1}^n$, so $v_n \in \XSet_n = \spn \{ \ell_1, \ldots, \ell_n \}$.
@@
 The same can be said about \cref{eq:RBFInterpolation}, which defines another choice
-of the interpolant. Another choice of projector $P_n$.
+of the interpolant, and another choice of projector $P_n$. A fundamental property of
+interpolatory projectors is the following (see \cite[Proposition
+4.1]{Avitabile:2023ab}): \da[]{Cite Atkinson and others} for any
+$r \in \XSet$, 
+\begin{equation}\label{eq:projCollProperty}
+P_n r = 0, 
+\qquad \text{if, and only if,} \qquad 
+(P_n r)(x_i) = 0, \qquad i \in \NSet_n.
+\end{equation}
+This property is useful in describing the system in terms of residual, as we shall do
+in a moment.
 
-Writing things in terms of $P_n$ is useful, as it compactifies the notation. Firstly,
-it is possible to rewrite the neural field equation \cref{eq:NF} as the following
-Cauchy problem on the Banach space $\XSet$
+We now revisit the neural field equation and the collocation scheme in terms of the
+projector $P_n$. Firstly, we rewrite the neural field equation \cref{eq:NF} in
+operator form, as the following Cauchy problem on the Banach space $\XSet$
 \begin{equation}\label{eq:cauchy}
   \begin{aligned}
-  & \dot U(t) = - U(t) + WF(U(t)sketch.pnsketch.pngg) + G(t) =:  N(U(t),t), \qquad t \in J, \\
+  & U'(t) = - U(t) + WF(U(t)) + G(t) =:  N(t,U(t)), \qquad t \in J, \\
   & U(0) = V.
@@
 \end{equation}
-The identification with \cref{eq:NF} is established with the positions
+The identification with \cref{eq:NF} is established by setting $J = [0,T]$,
+$V = v(\blank)$, and
 \[  
   \begin{aligned}
-   & V = v(\blank)\\
-   & U \colon [0,T] \to \XSet, && t \mapsto u(\blank,t) \\
-   & G \colon [0,T] \to \XSet, && t \mapsto g(\blank,t) \\
-   & W \colon \XSet \to \XSet, && v \mapsto \int_{D} w(\blank,y) v(y) \,d \mu(x') \\
+   & U \colon J \to \XSet, && t \mapsto u(\blank,t),
+   & G \colon J \to \XSet, && t \mapsto g(\blank,t), \\
+   & W \colon \XSet \to \XSet, && v \mapsto \int_{D} w(\blank,y) v(y) \,d \mu(y),
    & F \colon \XSet \to \XSet, && v \mapsto f(v). 
@@
 \]
-The Cauchy problem is well-posed under general assumptions that is, there is a unique
-$u \in C^1(J,\XSet)$ satisfying \cref{eq:cauchy}.
+Under suitable generic assumptions \da[]{if needed, include them}, the Cauchy problem
+is well-posed under general assumptions that is, there is a unique $u \in
+C^1(J,\XSet)$ satisfying \cref{eq:cauchy}, see
+\cite{Potthast:2009bd,Faugeras:2009gn}, and \cite[Lemma 2.7]{Avitabile:2023ab}.
 
@@
   \begin{aligned}
-  & \dot U_n(t) = - U_n(t) + P_n WF(U(t)) + P_n G(t) =  P_n N(U_n(t),t), \qquad t \in J, \\
+  & \dot U_n(t) = - U_n(t) + P_n WF(U_n(t)) + P_n G(t) =  P_n N(t,U_n(t)), \qquad t \in J, \\
   & U_n(0) = P_n V,
@@
 \end{equation}
-which is a Cauchy problem on the space $\XSet_{n}$.
+which is a Cauchy problem on the space $\XSet_{n}$. This problem is also well-posed
+and gives a solution $u_n \in C^1(J,\XSet_n)$ and
+\cite[Theorem 3.1]{Avitabile:2023ab}.
+
+Using \cref{eq:projCollProperty}, we can now pass from \cref{eq:cauchyPn} to an
+equivalent ODE in $\RSet^n$. We define the residual
+\[
+  R \colon C^1(J,\XSet) \to C(J,\XSet), \qquad U \mapsto U' - N(\blank, U),
+\]
+rewrite \cref{eq:cauchyPn} as
+\[
+  P_n \bigl( R(U_n)(t) \bigr) = 0, \qquad t \in J, \qquad 
+  P_n \bigl( U_n(0) - V\bigr) = 0, 
+\]
+which, by property \cref{eq:projCollProperty} is equivalent to
+\[
+  \begin{aligned}
+  & U'_n(t)(x_i) = \bigl[P_n N(t,U_n(t))\bigr](x_i), \qquad (i,t) \in \NSet_n \times J, \\
+  & U_n(0)(x_i) = (P_n V)(x_i).
+  \end{aligned}
+\]
+Finally, since $U_n(t) \in \XSet_n$, we can write $U_n(t) = \sum_{j \in \NSet_n}
+\alpha_j(t) \ell_j$ and obtain the sought ODE on $\RSet^n$,
+\begin{equation} \label{eq:NFODE}
+  \begin{aligned}
+   & \alpha'(t) = -\alpha(t) + K(\alpha(t)) + \gamma(t) \qquad  t \in [0,T],\\ 
+   & \alpha (0) = \nu,
+  \end{aligned}
+\end{equation}  
+where 
+\[
+  \begin{aligned}
+  & \alpha_i(t) = U_n(t)(x_i), && \nu_i = v(x_i),  \\
+  & \gamma_i(t) = g(x_i,t),
+  && K_i(\alpha) = 
+    \int_{D}w(x_i,y) f\Bigl( \sum_{j \in \NSet_n} \alpha_j \ell_j(y)\Bigr) \, d\mu(y)
+  \end{aligned}
+\]
+In passing, we note that the system \cref{eq:NFODE} above is related, but not
+identical, to \cref{eq:NFODEQuad}. The former is expressed in terms of 
+$K \colon \RSet^n \to \RSet^n$, the latter in terms of
+$\tilde K \colon \RSet^n \to \RSet^n$, which adopts a quadrature rule to approximate
+the integrals in $K$.
 
@@
 
-\subsection{References} \label{sec:references}
-These are random citations \cite{KoMa14,siam}. For cross-references I use
+\subsection{References} \label{sec:references} For cross-references I use
 \texttt{cleveref}: like this

### 2023-11-14 Manuscript/manuscript.tex
First pass at explaining the algorithm

@@
 Finally, since $U_n(t) \in \XSet_n$, we can write $U_n(t) = \sum_{j \in \NSet_n}
-\alpha_j(t) \ell_j$ and obtain the sought ODE on $\RSet^n$,
+\alpha_j(t) \ell_j$ and an ODE on $\RSet^n$,
 \begin{equation} \label{eq:NFODE}
@@
 $\tilde K \colon \RSet^n \to \RSet^n$, which adopts a quadrature rule to approximate
-the integrals in $K$.
+the integrals in $K$. In the next section, we will give a concrete example of $\tilde
+K$
 
@@
   \label{fig:sketch}
+  \da[inline]{We need a meaningful sketch here}
 \end{figure}
 
-\section{Conventions} \label{sec:conventions}
-See \texttt{docsiamart.pdf} for many examples. I generally adopt the conventions of
-that template. I remark a few things below.
+We now introduce the interpolant we will use in the paper. To define the interpolant
+we start from the set of $n$ nodes $\Xi = \{ x_i \}_{i \in \NSet_n}$, with the
+associated \textit{Voronoi diagram} $\{ D_i \}_{i \in \NSet_n}$, that is, the
+partition of $D$ into cells
+\[
+  D_i = \Bigl\{ x \in D \colon \| x - x_i \| = \min_{j \in \NSet_n} \| x - x_j \| \Bigr\}.
+\]
+Further, we consider a map associating each node $x_i$ to $k_i$ nodes $\{\xi_{ij}
+\}_{j = 1}^{k_i}$ in $\Xi$, the stencil of $x_i$. More formally we define the
+mapping
+\[
+  S \colon \Xi \to 2^{\,\Xi}, \qquad x_i \mapsto \{ \xi_{ij} \}_{j = 1}^{k_i},
+\]
+where $2^{\,\Xi}$ is the power set $\Xi$, and denote by $S(x_i)$, or simply $S_i$, the
+stencil of node $i$. We stipulate that $x_i$ is part of its stencil, $x_i \in
+S(x_i)$. Finally, we define a basic radial function $\phi = \Phi( \| \blank \|)
+\colon \RSet^d \to \RSet$, and construct an interpolant of $v$ in the
+following form
+\begin{equation}\label{eq:RBFExpStencil}   
+  v_n(x) = \sum_{j = 1}^{n} \sum_{l=1}^{k_j} V_{jl} \Phi(\| x - \xi_{jl} \|) 1_{D_j}(x) 
+\end{equation}
+where $1_A$ denotes the indicator function of the set $A \subset \RSet^d$. In
+\cref{eq:RBFExpStencil}, the interpolant sums over the stencil in each node, and it
+is therefore expressed in terms of $k_1 + \dots + k_n \geq n$ coefficients $\{
+V_{jl}\}$. Since $v_n$ interpolates $v$ at the nodes $\Xi$, the coefficients
+$V_{jl}$ satisfy the $n$ equations
+\[
+  v(x_j) = \sum_{l=1}^{k_j} V_{jl} \Phi(\| x_j - \xi_{jl} \|), \qquad j \in \NSet_n.
+\]
+The system above does not uniquely specify the coefficients $V_{jl}$, and so further
+constraints are added,
+\da[inline]{Complete this after Sage's input}
 
-\subsection{References} \label{sec:references} For cross-references I use
-\texttt{cleveref}: like this
-\cref{eq:test}, like this \Cref{eq:test}, or like this \cref{sec:conventions}, or
-this \cref{hyp:trivial}, or this \cref{clm:trivial}.
-\begin{equation}\label{eq:test}
-  x = y
-\end{equation}
+Crucially, it is possible to pass from the form \cref{eq:RBFExpStencil} of the
+interpolant, to the so-called \textit{cardinal form} of the interpolant, that, is a
+Lagrange interpolant
+\[
+  v_n(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x),
+\]
+using the transformation
+\da[inline]{Complete after Sage's input}
 
-\subsection{Todos}
-You can find templates of other author's todos in
-\texttt{damacros.sty}, which you are welcome to edit.
-This is how I make a short comment\da{Short comment by Daniele}. Comments can also go
-inline like the one below.
-\da[inline]{Long comments by Daniele} 
-% \missingfigure
-% 
-% \section*{Acknowledgments}
+Henceforth, we will use the cardinal form for the analysis and the presentation of
+our scheme, while we resort to the stencil-based form to carry out concrete
+calculations. The cardinal form of the interpolant defines a projector $P_n \colon \XSet \to
+\XSet_n = \spn \{ \ell_1, \ldots, \ell_n \}$, and therefore fits into the framework
+presented in the past section.
 
-\section{Testing commands}
-
-\subsection{Testing maths operators}
-Here are tests for operators
+To complete the scheme, we now require a quadrature formula for the integrand. The
+formula we employ is a composite formula over the Voronoi diagram
 \[
-  \spec, \quad 
-  \range, \quad 
-  \spn, \quad
-  \diag, \quad
-  \meas, \quad
-  \id, \quad
-  \real z, \quad
-  \imag z, \quad
-  \e, \quad
-  \iunit, \quad
+  \begin{aligned}
+  Q v := \int_{D} v(y)\,d \mu(y) 
+    & = \sum_{j=1}^n \int_{D_j} v(y) \,d \mu(y) \\
+    & \approx \sum_{j = 1}^n \int_{D_j} v_n(y) \,d \mu(y)
+      \approx \sum_{j = 1}^n v(x_j) \mu_j =: Q_n v,
+  \end{aligned}
 \]
-
-\subsection{Testing additional theorem-like environments} The following additional
-theorem-like environments may be useful.
-
-\begin{hypothesis} \label{hyp:trivial}
-  This is a hypothesis
-\end{hypothesis}
-
-\begin{claim} \label{clm:trivial}
-  This is a claim
-\end{claim}
-
-\subsection{Testing sets}
-This are a few examples of sets (see more commands in \texttt{damacros.sty})
+where the weights $\{ \mu_j \}_{j \in \NSet_n}$ satisfy
+\da[inline]{Complete. I am unsure there are two approximation symbols above, or just
+one. If possible, say something about the expected convergence rate of $|Qv - Q_nv|$.} 
+We are now ready to specify a discrete scheme by applying the quadrature rule $Q_n$
+the function $K$ in \cref{eq:NFODE}, recalling that each $D_i$ contains exactly
+one node, namely $x_i$, and obtaining the spatially-discrete scheme
 \[
-  \RSet, \quad
-  \CSet, \quad
-  \ZSet, \quad
-  \sph, \quad
-  \ldots
+  \begin{aligned}
+  & \tilde \alpha'(t) = -\tilde \alpha(t) + \gamma(t) 
+    + M f(\tilde \alpha(t)), \qquad t \in J \\
+  & \tilde \alpha(0) = \nu
+  \end{aligned}
 \]
-
-\subsection{Empty arguments}
-Command for empty arguments, which require spacing
-\[
-  u(\blank,t) \quad t \in \RSet
-\]
-
-\subsection{Calligraphic fonts}
-If one really must use calligraphic fonts, there are macros for this.
-\[
-  \calA, \quad
-  \calL, \quad
-  \calM, \quad
-  \calS, \quad
-\]
+where $M_{ij} = w(x_i,x_j)\mu_j$, and in which we have used, with a little abuse of
+notation, $f(\alpha)$ to denote the vector with components $(f(\alpha))_i =
+f(\alpha_i)$. We have therefore provided an RBF example of the collocation scheme
+\cref{eq:NFODEQuad}, in which $\tilde K(\tilde \alpha) = M f(\tilde \alpha)$.
 

### 2023-11-14 Manuscript/manuscript.tex
Sentence about the quadrature rivisited

@@
 \]
-where the weights $\{ \mu_j \}_{j \in \NSet_n}$ satisfy
+in which the integral over a cell $D_i$ is approximated using a single node $x_j$,
+the centre of the cell, and a single weight $\mu_j$. The weights satisfy \ldots
 \da[inline]{Complete. I am unsure there are two approximation symbols above, or just
-one. If possible, say something about the expected convergence rate of $|Qv - Q_nv|$.} 
-We are now ready to specify a discrete scheme by applying the quadrature rule $Q_n$
+one. That depends on how weights are calculated. If possible, we should say something
+about the expected convergence rate of $|Qv - Q_nv|$.} We are now ready to specify a discrete scheme by applying the quadrature rule $Q_n$
 the function $K$ in \cref{eq:NFODE}, recalling that each $D_i$ contains exactly

### 2023-11-14 Manuscript/manuscript.tex
Small changes

@@
 \]
-that is, $v_n$ is the action of $P_n$ on the function $v$, $v_n = P_n v$.
-The same can be said about \cref{eq:RBFInterpolation}, which defines another choice
-of the interpolant, and another choice of projector $P_n$. A fundamental property of
-interpolatory projectors is the following (see \cite[Proposition
-4.1]{Avitabile:2023ab}): \da[]{Cite Atkinson and others} for any
-$r \in \XSet$, 
+that is, $v_n$ is the action of $P_n$ on the function $v$, $v_n = P_n v$. As we shall
+see in the next section, the same can be said about \cref{eq:RBFInterpolation}, which
+defines another choice of the interpolant, and another choice of projector $P_n$. A
+useful property of interpolatory projectors is the following (see
+\cite{atkinson1997,atkinson2005theoretical} and \cite[Proposition
+4.1]{Avitabile:2023ab}): for all $r \in \XSet$
 \begin{equation}\label{eq:projCollProperty}
@@
 \end{equation}
-This property is useful in describing the system in terms of residual, as we shall do
+This property is useful in describing the system in terms of residuals, as we shall do
 in a moment.
@@
 which is a Cauchy problem on the space $\XSet_{n}$. This problem is also well-posed
-and gives a solution $u_n \in C^1(J,\XSet_n)$ and
-\cite[Theorem 3.1]{Avitabile:2023ab}.
+\cite[Theorem 3.1]{Avitabile:2023ab}, and gives a solution $u_n \in C^1(J,\XSet_n)$.
 
@@
 the integrals in $K$. In the next section, we will give a concrete example of $\tilde
-K$
+K$.
 
@@
 we start from the set of $n$ nodes $\Xi = \{ x_i \}_{i \in \NSet_n}$, with the
-associated \textit{Voronoi diagram} $\{ D_i \}_{i \in \NSet_n}$, that is, the
+associated \textit{Voronoi diagram} $\{ D_i \}_{i \in \NSet_n}$, that is,
+the\da[]{Perhaps it's ''a" not "the". I recall Sage commenting on $D$ not being
+uniquely defined if $D$ has a boundary?}
 partition of $D$ into cells
@@
 \]
-Further, we consider a map associating each node $x_i$ to $k_i$ nodes $\{\xi_{ij}
-\}_{j = 1}^{k_i}$ in $\Xi$, the stencil of $x_i$. More formally we define the
-mapping
+Further, we consider a map associating each node $x_i$ to its \textit{stencil}, that
+is, a set of $k_i$ nodes $\{\xi_{ij} \}_{j = 1}^{k_i}$ in $\Xi$. This is formalised
+by a mapping
 \[
@@
 \]
-where $2^{\,\Xi}$ is the power set $\Xi$, and denote by $S(x_i)$, or simply $S_i$, the
-stencil of node $i$. We stipulate that $x_i$ is part of its stencil, $x_i \in
-S(x_i)$. Finally, we define a basic radial function $\phi = \Phi( \| \blank \|)
-\colon \RSet^d \to \RSet$, and construct an interpolant of $v$ in the
-following form
+where $2^{\,\Xi}$ is the power set of $\Xi$. We denote by $S(x_i)$, or simply $S_i$,
+the stencil of node $i$, and we stipulate that $S(x_i)$ contains $x_i$. Finally, we
+pick a basic radial function $\phi = \Phi( \| \blank \|) \colon \RSet^d \to \RSet$,
+and seek for an interpolant of $v$ in the following form
 \begin{equation}\label{eq:RBFExpStencil}   
@@
 where $1_A$ denotes the indicator function of the set $A \subset \RSet^d$. In
-\cref{eq:RBFExpStencil}, the interpolant sums over the stencil in each node, and it
+\cref{eq:RBFExpStencil}, the interpolant sums over the stencil in each node, and 
 is therefore expressed in terms of $k_1 + \dots + k_n \geq n$ coefficients $\{
@@
 The system above does not uniquely specify the coefficients $V_{jl}$, and so further
-constraints are added,
+constraints are added. In summary, the $\{ V_{jl} \}$ is obtained by solving the
+following 
 \da[inline]{Complete this after Sage's input}
@@
 \da[inline]{Complete. I am unsure there are two approximation symbols above, or just
-one. That depends on how weights are calculated. If possible, we should say something
-about the expected convergence rate of $|Qv - Q_nv|$.} We are now ready to specify a discrete scheme by applying the quadrature rule $Q_n$
-the function $K$ in \cref{eq:NFODE}, recalling that each $D_i$ contains exactly
-one node, namely $x_i$, and obtaining the spatially-discrete scheme
+one. That depends on how weights are calculated, and on the function space of $v$, I
+believe. If possible, we should say something
+about the expected convergence rate of $|Qv - Q_nv|$.} We are now ready to specify a
+discrete scheme by applying the quadrature rule $Q_n$ to each of the elements $K_i$ in
+\cref{eq:NFODE}. Recalling that each $D_i$ contains exactly one node, namely $x_i$,
+we obtain the spatially-discrete scheme
 \[
   \begin{aligned}
-  & \tilde \alpha'(t) = -\tilde \alpha(t) + \gamma(t) 
-    + M f(\tilde \alpha(t)), \qquad t \in J \\
+  & \tilde \alpha'(t) = -\tilde \alpha(t) 
+    + M f(\tilde \alpha(t))+ \gamma(t) , \qquad t \in J \\
   & \tilde \alpha(0) = \nu
@@
 where $M_{ij} = w(x_i,x_j)\mu_j$, and in which we have used, with a little abuse of
-notation, $f(\alpha)$ to denote the vector with components $(f(\alpha))_i =
+notation, $f(\alpha)$ to denote the vector with components $f_i(\alpha) =
 f(\alpha_i)$. We have therefore provided an RBF example of the collocation scheme

### 2023-11-14 Simulations/driver.py
Moved code from old repo.

@@ -1,0 +1,132 @@
+from functools import partial
+import matplotlib
+import matplotlib.pyplot as plt
+
+import numpy as np
+import numpy.linalg as la
+
+from rbf.rbf import PHS
+from rbf.points import UnitSquare
+from rbf.interpolate import interpolate, LocalInterpolator
+
+from scipy.special import roots_chebyt as cheb
+from tqdm import tqdm
+
+
+########################
+# two test functions
+########################
+def Frankes_function(x, y):
+    """A common function to test multidimensional interpolation."""
+    return (
+        0.75 * np.exp(-1 / 4 * ((9 * x - 2) ** 2 + (9 * y - 2) ** 2))
+        + 0.75 * np.exp(-1 / 49 * ((9 * x + 12) ** 2 + (9 * y + 1) ** 2))
+        + 0.5 * np.exp(-1 / 4 * ((9 * x - 7) ** 2 + (9 * y - 3) ** 2))
+        + 0.2 * np.exp(-((9 * x - 4) ** 2 + (9 * y - 7) ** 2))
+    )
+
+
+def poly_test(x, y):
+    """A polynomial test function."""
+    return 4 - 3 * x + 2 * y - x**2 + x * y - y**2
+
+
+# test_func = poly_test
+test_func = Frankes_function
+
+
+########################
+# Dense sample points
+########################
+xs_dense, ys_dense = np.meshgrid(np.linspace(0, 1, 401), np.linspace(0, 1, 401))
+points_dense = np.block([[xs_dense.ravel()], [ys_dense.ravel()]]).T
+fs_dense = test_func(xs_dense, ys_dense)
+
+########################
+# RBF sample points
+########################
+
+xs, ys = UnitSquare(21, verbose=True).points.T
+
+# Cartesian points
+# xs, ys = np.meshgrid(
+#     np.linspace(np.min(xs_dense), np.max(xs_dense), 11),
+#     np.linspace(np.min(ys_dense), np.max(ys_dense), 11),
+# )
+
+
+# Chebshev points
+# def cheb_nodes_1D(n, x_min, x_max):
+#     points = (cheb(n)[0] + 1)/2
+#     points = (x_max - x_min)*points + x_min
+#     return points
+#
+#
+# xs, ys = np.meshgrid(
+#         cheb_nodes_1D(15, np.min(xs_dense), np.max(xs_dense)),
[…]

### 2023-11-14 Simulations/rbf/interpolate.py

@@ -1,0 +1,100 @@
+"""
+An RBF interpolant class.
+"""
+
+import numpy as np
+import numpy.linalg as la
+from .poly_utils import poly_powers
+from .rbf import RBF, PHS, pairwise_dist
+from scipy.spatial import KDTree
+from .stencil import Stencil
+
+
+class Interpolator:
+    def __init__(
+        self, *, stencil: Stencil, fs: np.ndarray, rbf: RBF, poly_deg: int = -1
+    ):
+        assert len(fs) == stencil.num_points
+        self.stencil = stencil
+        self.rbf = rbf
+        self.poly_deg = poly_deg
+        self.find_weights(fs)
+
+    def find_weights(self, fs: np.ndarray) -> tuple[np.ndarray, np.ndarray]:
+        A = self.stencil.interpolation_matrix(rbf=self.rbf, poly_deg=self.poly_deg)
+        ys = np.zeros(A.shape[0])
+        ys[: self.stencil.num_points] = fs
+        try:
+            weights = la.solve(A, ys)
+        except la.LinAlgError:
+            raise ValueError(
+                """Interpolation matrix is singular.
+                I haven't yet implemented the Schur decomp solve yet.
+                For now, use shape parameter RBFs."""
+            )
+        self.rbf_weights = weights[: self.stencil.num_points]
+        self.poly_weights = weights[self.stencil.num_points :]
+
+    def __call__(self, z):
+        z = self.stencil.shift_points(z)
+        rbf_val = sum(
+            w * self.rbf(la.norm(z - point))
+            for w, point in zip(self.rbf_weights, self.stencil.scaled_points)
+        )
+        poly_val = sum(
+            w * poly(z)
+            for w, poly in zip(
+                self.poly_weights,
+                poly_powers(dim=self.stencil.dim, max_deg=self.poly_deg),
+            )
+        )
+        return rbf_val + poly_val
+
+
+class LocalInterpolator:
+    def __init__(
+        self,
+        *,
+        points: np.ndarray[float],
+        fs: np.ndarray[float],
+        rbf: RBF,
+        poly_deg: int,
+        stencil_size: int
+    ):
+        self.points = points
+        self.fs = fs
+        self.rbf = rbf
+        self.poly_deg = poly_deg
+        self.stencil_size = stencil_size
+
+        self.kdt = KDTree(points)
+        self.generate_stencils()
+        self.interpolate_on_stencils()
+
+    def generate_stencils(self):
+        self.stencils = []
[…]

### 2023-11-14 Simulations/rbf/points/unit_square.py

@@ -1,0 +1,89 @@
+import matplotlib.pyplot as plt
+import numpy as np
+from .points import PointCloud, GaussianRepulsionKernel
+from tqdm import tqdm
+
+
+# make a square with N x N points
+
+
+class UnitSquare(PointCloud):
+    """
+    Generates N^2 points in unit square [0, 1] x [0, 1] with N points along each
+    boundary.
+    """
+
+    def __init__(self, N: int, auto_settle=True, verbose=False):
+        self.N = N
+        xs, ys = np.meshgrid(*2 * (np.linspace(0, 1, N),))
+        points = np.block([[xs.ravel()], [ys.ravel()]]).T
+        boundary_mask = (
+            (points[:, 0] == 0.0)
+            | (points[:, 0] == 1.0)
+            | (points[:, 1] == 0.0)
+            | (points[:, 1] == 1.0)
+        )
+
+        num_boundary = (N - 1) * 4
+        sorted_points = np.empty_like(points)
+        sorted_points[:-num_boundary] = points[~boundary_mask]
+        sorted_points[-num_boundary:] = points[boundary_mask]
+        super(UnitSquare, self).__init__(
+            sorted_points, num_interior=N**2 - num_boundary, num_boundary=num_boundary
+        )
+        if auto_settle:
+            self.auto_settle(verbose=verbose)
+
+    def force_shape(self, x):
+        return (1 + np.tanh(-self.N*2 * x)) / 2
+
+    def boundary_force(self, point):
+        force = np.zeros_like(point)
+        x, y = point
+        force[0] = self.force_shape(x) - self.force_shape(1 - x)
+        force[1] = self.force_shape(y) - self.force_shape(1 - y)
+        return 4 * force
+
+    def settle(self, rate: float, repeat: int = 1, verbose=False):
+        kernel = GaussianRepulsionKernel(height=4, shape=2 / self.N)
+        # rate = 0.2 / N
+        # radius = 3.1 / N
+        num_neighbors = 18
+        my_iter = range(repeat)
+        if verbose:
+            my_iter = tqdm(my_iter)
+        for _ in my_iter:
+            super(UnitSquare, self).settle(
+                kernel=kernel,
+                rate=rate / self.N,
+                num_neighbors=num_neighbors,
+                force=self.boundary_force,
+            )
+
+    def auto_settle(self, verbose=False):
+        for rate, repeat in [
[…]

### 2023-11-14 Simulations/rbf/poly_utils.py

@@ -1,0 +1,94 @@
+"""
+Utilities for multidimensional polynomials.
+
+RBF methods can included polynomial basis terms to increase accuracy.
+Polynomial basis terms are also necessary to ensure invertability of
+RBF matrices based on polyharmonic spline RBFs.
+"""
+
+from functools import cache, reduce
+from math import comb, perm
+import numpy as np
+import operator
+from typing import Collection, Generator
+
+
+class Monomial:
+    def __init__(self, *pows: tuple[int]):
+        self.pows = pows
+
+    def __call__(self, points: np.ndarray[float]):
+        """
+        Apply the polynomial powers to an array of points, pointwise.
+        """
+        shape = points.shape
+        if len(self.pows) == 1:
+            points = points.reshape(*shape, 1)
+        ret = reduce(
+            operator.mul,
+            (x**p for x, p in zip(points.T, self.pows)),
+        )
+        if len(self.pows) == 1:
+            return ret.flatten()
+        return ret
+
+    def diff(self, points: np.ndarray[float], *orders: Collection[int]):
+        shape = points.shape
+        if len(self.pows) == 1:
+            points = points.reshape(*shape, 1)
+        new_pows = []
+        coeff = 1
+        for o, p in zip(orders, self.pows):
+            new_pows.append(p - o)
+            coeff *= perm(p, o)
+        ret = coeff * reduce(
+            operator.mul,
+            (x**max(p, 0) for x, p in zip(points.T, new_pows)),
+        )
+        if len(self.pows) == 1:
+            return ret.flatten()
+        return ret
+
+
+PolyPowGen = Generator[Monomial, None, None]
+
+
+@cache
+def poly_basis_dim(dim: int, deg: int) -> int:
+    """
+    The number of polynomial basis terms up to a given degree for a given
+    dimension.
+    """
+    return comb(dim + deg, deg)
+
+
+def poly_powers_of_deg(dim: int, deg: int) -> PolyPowGen:
+    """
+    Find all combinations of exponents of polynomials of the specified degree for
+    the specified dimension. Return a generator that yields tuples of the exponents.
+    """
+    if dim == 1:
+        yield Monomial(deg)
+    else:
+        for d in range(deg + 1):
+            yield from (
+                Monomial(*poly.pows, d) for poly in poly_powers_of_deg(dim - 1, deg - d)
+            )
+
+
+def poly_powers_gen(dim: int, max_deg: int) -> PolyPowGen:
+    """
[…]

### 2023-11-14 Simulations/rbf/rbf.py

@@ -1,0 +1,79 @@
+"""
+Classes for radial basis functions.
+"""
+
+from abc import ABC, abstractmethod, abstractstaticmethod
+import numpy as np
+import numpy.linalg as la
+
+
+def pairwise_diff(p1: np.ndarray[float], p2: np.ndarray[float]) -> np.ndarray[float]:
+    """Return a matrix of the poirwise differences between two
+    vectors of points.
+    The points must be of the same dimension.
+    """
+    return np.array([[(x - y) for x in p1] for y in p2])
+
+
+def pairwise_dist(p1: np.ndarray[float], p2: np.ndarray[float]) -> np.ndarray[float]:
+    """Return a matrix of the pairwise distances between two
+    vectors of points.
+    The points must be of the same dimension.
+    """
+    return np.array([[la.norm(x - y) for x in p1] for y in p2])
+
+
+class RBF(ABC):
+    @abstractmethod
+    def __call__(r):
+        raise NotImplementedError
+
+    @abstractmethod
+    def dr(r):
+        raise NotImplementedError
+
+    @abstractmethod
+    def dr_div_r(r):
+        raise NotImplementedError
+
+
+class OddPHS(RBF):
+    def __init__(self, deg: int):
+        assert type(deg) is int
+        assert deg > 0
+        assert deg % 2 == 1
+        self.deg = deg
+
+    def __call__(self, r):
+        return r**self.deg
+
+    def dr(self, r):
+        return self.deg * r ** (self.deg - 1)
+
+    def dr_div_r(self, r):
+        return self.deg * r ** (self.deg - 2)
+
+
+class EvenPHS(RBF):
+    def __init__(self, deg: int):
+        assert type(deg) is int
+        assert deg > 0
+        self.deg = deg
+        self.even = deg % 2 == 0
+
+    def __call__(self, r):
+        # account for removeable singularity at r = 0
+        ret = np.empty(r.shape)
+        mask = ret == 0
+        ret[mask] = 0
+        ret[~mask] = r[~mask]**self.deg * np.log(r[~mask])
+        return ret
+
+    def dr(self, r):
+        raise NotImplementedError
+
+
+def PHS(deg: int) -> RBF:
+    if deg % 2 == 0:
+        return EvenPHS(deg)
+    return OddPHS(deg)

### 2023-11-14 Simulations/rbf/stencil.py

@@ -1,0 +1,66 @@
+"""
+Stencil class for RBF interpolation.
+"""
+
+from functools import cache
+import numpy as np
+import numpy.linalg as la
+from .poly_utils import poly_basis_dim, poly_powers
+from .rbf import RBF, pairwise_diff, pairwise_dist
+
+
+class Stencil:
+    def __init__(
+        self,
+        points: np.ndarray[float],
+    ):
+        shape = points.shape
+        if len(shape) == 1:
+            self.dim = 1
+        else:
+            assert len(shape) == 2
+            self.dim = shape[1]
+        self.points = points
+
+        self.upper_bounds = np.max(points, axis=0)
+        self.lower_bounds = np.min(points, axis=0)
+
+    @property
+    @cache
+    def num_points(self):
+        return len(self.points)
+
+    def shift_points(self, points: np.ndarray) -> np.ndarray:
+        mids = (self.upper_bounds + self.lower_bounds) / 2
+        widths = self.upper_bounds - self.lower_bounds
+        return (points - mids) / widths
+
+    @property
+    @cache
+    def scaled_points(self):
+        return self.shift_points(self.points)
+
+    @property
+    def pairwise_diff(self) -> np.ndarray[float]:
+        points = self.scaled_points
+        return pairwise_diff(points, points)
+
+    @property
+    def dist_mat(self) -> np.ndarray[float]:
+        points = self.scaled_points
+        return pairwise_dist(points, points)
+
+    def rbf_mat(self, rbf: RBF):
+        return rbf(self.dist_mat)
+
+    def poly_mat(self, poly_deg: int) -> np.ndarray[float]:
+        P = np.ones((self.num_points, poly_basis_dim(self.dim, poly_deg)))
+        for index, poly in enumerate(poly_powers(self.dim, max_deg=poly_deg)):
+            P[:, index] = poly(self.scaled_points)
+        return P
+
+    def interpolation_matrix(self, rbf: RBF, poly_deg: int = -1) -> np.ndarray[float]:
+        P = self.poly_mat(poly_deg=poly_deg)
+        basis_size = P.shape[1]
+        zeros = np.zeros((basis_size, basis_size))
+        return np.block([[self.rbf_mat(rbf=rbf), P], [P.T, zeros]])

### 2023-11-14 Simulations/drivers/fd_1D_driver.py
1D finite difference working

@@ -1,0 +1,51 @@
+import set_path
+
+import matplotlib.pyplot as plt
+import numpy as np
+from rbf import RBF, PHS
+from rbf.finite_differences import Derivative1D
+from rbf.linear_functional import FunctionalStencil, LinearFunctional
+from rbf.poly_utils import Monomial
+
+
+def foo(x):
+    return np.sin(x)
+
+
+def d_foo(x):
+    return np.cos(x)
+
+
+rbf = PHS(3)
+poly_deg = 4
+k = 5
+hs = [2 ** (-i) for i in range(20)]
+assert k % 2 == 1
+
+print("Function: sin(x)")
+print(f"1st derivative evaluated at 0: {d_foo(0)}")
+print("Stencil:")
+unshifted_points = np.concatenate([np.arange(-k // 2 + 1, 0), np.arange((k + 1) // 2)])
+print(unshifted_points)
+
+errors = []
+for h in hs:
+    points = unshifted_points * h
+    stencil = FunctionalStencil(points)
+    weights = stencil.weights(rbf, Derivative1D(), poly_deg)
+    error = abs(weights @ foo(points) - d_foo(0))
+    errors.append(error)
+
+plt.loglog(hs, errors, 'k.-')
+plt.xlabel("h")
+plt.ylabel("Error")
+plt.title(f"Finite Difference Test:\nrbf: {rbf}, poly degree: {poly_deg}, stencil size: {k}")
+
+start = 1
+end = 1e-3
+
+for order in range(1, 5):
+    plt.plot([start, end], [errors[0], end**order], label=f"$\\mathcal{{O}}({order})$")
+
+plt.legend()
+plt.show()

### 2023-11-14 Simulations/drivers/interpolation_driver.py

@@ -1,3 +1,9 @@
+"""
+A driver testing 2D interpolation on scattered nodes.
+"""
+
+import set_path
+
 from functools import partial
 import matplotlib
 import matplotlib.pyplot as plt

### 2023-11-14 Simulations/drivers/set_path.py

@@ -1,0 +1,11 @@
+"""
+We assume that the rbf module is in the parent directory of these drivers.
+We will add the parent directory to the path.
+"""
+
+from os.path import join, dirname, realpath
+import sys
+
+rbf_path = join(dirname(dirname(realpath(__file__))))
+sys.path.append(rbf_path)
+print(f"Added to PYTHONPATH: {rbf_path}")

### 2023-11-14 Simulations/rbf/finite_differences.py

@@ -1,0 +1,34 @@
+
+from .rbf import RBF, PHS
+from .poly_utils import Monomial
+from .linear_functional import LinearFunctional
+
+import numpy as np
+
+
+class Derivative1D(LinearFunctional):
+    def rbf_op(self, rbf: RBF, r: float, d: np.ndarray[float]) -> float:
+        return d * rbf.dr_div_r(r)
+
+    def poly_op(self, poly: Monomial, d: np.ndarray[float]) -> float:
+        return poly.diff(d, 1)
+
+    @property
+    def scale_powers(self) -> np.ndarray[int]:
+        return np.array([1])
+
+
+class Laplacian(LinearFunctional):
+
+    def __init__(self, dim: int):
+        self.dim = dim
+
+    def rbf_op(self, rbf: RBF, r: float, d: np.ndarray[float]) -> float:
+        return rbf.d2r(r) + (self.dim-1)*rbf.dr_div_r(r)
+
+    def poly_op(self, poly: Monomial, d: np.ndarray[float]) -> float:
+        return poly.diff(d, *[2 for _ in range(self.dim)])
+
+    @property
+    def scale_powers(self) -> np.ndarray[int]:
+        return np.array([2 for _ in range(self.dim)])

### 2023-11-14 Simulations/rbf/linear_functional.py

@@ -1,13 +1,20 @@
-from abc import ABC, abstractmethod
+from abc import ABC, abstractmethod, abstractproperty
+from functools import reduce
 import numpy as np
 import numpy.linalg as la
+from operator import mul
 from .poly_utils import Monomial, poly_basis_dim, poly_powers_gen
 from .rbf import RBF, pairwise_dist, pairwise_diff
 from .stencil import Stencil
 from typing import Callable
 
 
-class LinearOperator(ABC):
+class LinearFunctional(ABC):
+    """
+    A base class for linear functionals.
+    Be sure to account for stencil scaling.
+    """
+
     @abstractmethod
     def rbf_op(self, rbf: RBF, r: float, d: np.ndarray[float]) -> float:
         raise NotImplementedError
@@ -16,16 +23,20 @@ class LinearOperator(ABC):
     def poly_op(self, poly: Monomial, d: np.ndarray[float]) -> float:
         raise NotImplementedError
 
+    @abstractproperty
+    def scale_powers(self) -> np.ndarray[int]:
+        raise NotImplementedError
 
-class OperatorStencil(Stencil):
+
+class FunctionalStencil(Stencil):
     def __init__(
         self,
         points: np.ndarray[float],
     ):
-        super(OperatorStencil, self).__init__(points)
+        super(FunctionalStencil, self).__init__(points)
         self.center = points[0]
 
-    def weights(self, rbf: RBF, op: LinearOperator, poly_deg: int):
+    def weights(self, rbf: RBF, op: LinearFunctional, poly_deg: int):
         d = self.pairwise_diff[0]
         r = self.dist_mat[0]
         mat = self.interpolation_matrix(rbf, poly_deg)
@@ -38,4 +49,5 @@ class OperatorStencil(Stencil):
             ]
         ).T
         weights = la.solve(mat, rhs)
-        return weights[: len(self.points)]
+        scale = reduce(mul, (dx**p for dx, p in zip(self.scalings, op.scale_powers)))
+        return weights[: len(self.points)] / scale

### 2023-11-14 Simulations/rbf/rbf.py

@@ -33,7 +33,15 @@ class RBF(ABC):
         raise NotImplementedError
 
     @abstractmethod
+    def d2r(r):
+        raise NotImplementedError
+
+    @abstractmethod
     def dr_div_r(r):
+        raise NotImplementedError
+
+    @abstractmethod
+    def __repr__(self):
         raise NotImplementedError
 
 
@@ -50,8 +58,14 @@ class OddPHS(RBF):
     def dr(self, r):
         return self.deg * r ** (self.deg - 1)
 
+    def d2r(self, r):
+        return self.deg * (self.deg-1) * r ** (self.deg - 1)
+
     def dr_div_r(self, r):
         return self.deg * r ** (self.deg - 2)
+
+    def __repr__(self):
+        return f"r^{self.deg}"
 
 
 class EvenPHS(RBF):
@@ -71,6 +85,12 @@ class EvenPHS(RBF):
 
     def dr(self, r):
         raise NotImplementedError
+
+    def d2r(r):
+        raise NotImplementedError
+
+    def __repr__(self):
+        return f"r^{self.deg} log(r)"
 
 
 def PHS(deg: int) -> RBF:

### 2023-11-14 Simulations/rbf/stencil.py

@@ -22,8 +22,10 @@ class Stencil:
             self.dim = shape[1]
         self.points = points
 
-        self.upper_bounds = np.max(points, axis=0)
-        self.lower_bounds = np.min(points, axis=0)
+        self.upper_bounds = np.max(points, axis=0, keepdims=True).flatten()
+        self.lower_bounds = np.min(points, axis=0, keepdims=True).flatten()
+        self.mids = (self.upper_bounds + self.lower_bounds) / 2
+        self.scalings = (self.upper_bounds - self.lower_bounds)*2
 
     @property
     @cache
@@ -31,9 +33,7 @@ class Stencil:
         return len(self.points)
 
     def shift_points(self, points: np.ndarray) -> np.ndarray:
-        mids = (self.upper_bounds + self.lower_bounds) / 2
-        widths = self.upper_bounds - self.lower_bounds
-        return (points - mids) / widths
+        return (points - self.mids)/self.scalings
 
     @property
     @cache

### 2023-11-15 Simulations/drivers/fd_1D_driver.py
Laplacian FD working

@@ -22,7 +22,6 @@ k = 5
 hs = [2 ** (-i) for i in range(20)]
 assert k % 2 == 1
 
-print("Function: sin(x)")
 print(f"1st derivative evaluated at 0: {d_foo(0)}")
 print("Stencil:")
 unshifted_points = np.concatenate([np.arange(-k // 2 + 1, 0), np.arange((k + 1) // 2)])

### 2023-11-15 Simulations/drivers/fd_2D_driver.py

@@ -1,0 +1,79 @@
+import set_path
+
+from itertools import product
+import matplotlib.pyplot as plt
+import numpy as np
+import numpy.linalg as la
+from rbf import RBF, PHS
+from rbf.finite_differences import Laplacian
+from rbf.linear_functional import FunctionalStencil, LinearFunctional
+from rbf.poly_utils import Monomial
+import sympy as sym
+
+x_sym, y_sym = sym.symbols("x y")
+
+sym_foo = sym.cos(x_sym-1)*sym.cos(y_sym-3)
+
+
+foo = sym.lambdify((x_sym, y_sym), sym_foo)
+L_foo = sym.lambdify((x_sym, y_sym), sym_foo.diff(x_sym, 2) + sym_foo.diff(y_sym, 2))
+
+
+rbf = PHS(3)
+poly_deg = 2
+
+stencil_width = 3
+
+
+hs = [2**-i for i in range(2, 20)]
+
+assert stencil_width % 2 == 1
+unshifted_points = list(
+    product(*2 * (np.arange(-(stencil_width-1) // 2, (stencil_width+1) // 2),))
+)
+unshifted_points.sort(key=lambda point: la.norm(point))
+unshifted_points = np.array(unshifted_points, dtype=float)
+# unshifted_points += np.random.rand(*unshifted_points.shape)*1e-1
+# unshifted_points -= unshifted_points[0]
+
+print(f"Laplacian evaluated at 0, 0: {L_foo(0, 0)}")
+print("Stencil:")
+print(unshifted_points)
+
+stencil = FunctionalStencil(unshifted_points)
+weights = stencil.weights(rbf, Laplacian(dim=2), poly_deg)
+plt.figure("Stencil")
+for w, point in zip(weights, unshifted_points):
+    plt.plot(*point, 'k.')
+    plt.text(*point, str(np.round(w, 4)))
+plt.xlim(-stencil_width / 1.9, stencil_width/1.9)
+plt.ylim(-stencil_width / 1.9, stencil_width/1.9)
+plt.title("Stencil and Weights")
+plt.show()
+
+errors = []
+for h in hs:
+    points = unshifted_points * h
+    stencil = FunctionalStencil(points)
+    weights = stencil.weights(rbf, Laplacian(dim=2), poly_deg)
+    test = weights @ foo(*points.T)
+    error = abs(test - L_foo(0, 0))
[…]

### 2023-11-15 Simulations/rbf/finite_differences.py

@@ -1,4 +1,3 @@
-
 from .rbf import RBF, PHS
 from .poly_utils import Monomial
 from .linear_functional import LinearFunctional
@@ -14,21 +13,25 @@ class Derivative1D(LinearFunctional):
         return poly.diff(d, 1)
 
     @property
-    def scale_powers(self) -> np.ndarray[int]:
-        return np.array([1])
+    def scale_power(self) -> np.ndarray[int]:
+        return 1
 
 
 class Laplacian(LinearFunctional):
+    """Not working in higher than 1D yet."""
 
     def __init__(self, dim: int):
         self.dim = dim
 
     def rbf_op(self, rbf: RBF, r: float, d: np.ndarray[float]) -> float:
-        return rbf.d2r(r) + (self.dim-1)*rbf.dr_div_r(r)
+        return rbf.d2r(r) + (self.dim - 1) * rbf.dr_div_r(r)
 
     def poly_op(self, poly: Monomial, d: np.ndarray[float]) -> float:
-        return poly.diff(d, *[2 for _ in range(self.dim)])
+        return sum(
+            poly.diff(d, *[2 * (i == j) for i in range(self.dim)])
+            for j in range(self.dim)
+        )
 
     @property
-    def scale_powers(self) -> np.ndarray[int]:
-        return np.array([2 for _ in range(self.dim)])
+    def scale_power(self) -> np.ndarray[int]:
+        return 2

### 2023-11-15 Simulations/rbf/linear_functional.py

@@ -24,7 +24,7 @@ class LinearFunctional(ABC):
         raise NotImplementedError
 
     @abstractproperty
-    def scale_powers(self) -> np.ndarray[int]:
+    def scale_power(self) -> int:
         raise NotImplementedError
 
 
@@ -44,10 +44,9 @@ class FunctionalStencil(Stencil):
         rhs[: len(self.points)] = op.rbf_op(rbf, r, d).ravel()
         rhs[len(self.points) :] = np.array(
             [
-                op.poly_op(poly, np.array([0.0]))
+                op.poly_op(poly, np.zeros(self.dim))
                 for poly in poly_powers_gen(self.dim, poly_deg)
             ]
-        ).T
+        )
         weights = la.solve(mat, rhs)
-        scale = reduce(mul, (dx**p for dx, p in zip(self.scalings, op.scale_powers)))
-        return weights[: len(self.points)] / scale
+        return weights[: len(self.points)] / (self.scaling**op.scale_power)

### 2023-11-15 Simulations/rbf/poly_utils.py

@@ -49,6 +49,9 @@ class Monomial:
             return ret.flatten()
         return ret
 
+    def __repr__(self):
+        return f"Monmial{self.pows}"
+
 
 PolyPowGen = Generator[Monomial, None, None]
 

### 2023-11-15 Simulations/rbf/rbf.py

@@ -59,7 +59,7 @@ class OddPHS(RBF):
         return self.deg * r ** (self.deg - 1)
 
     def d2r(self, r):
-        return self.deg * (self.deg-1) * r ** (self.deg - 1)
+        return self.deg * (self.deg-1) * r ** (self.deg - 2)
 
     def dr_div_r(self, r):
         return self.deg * r ** (self.deg - 2)

### 2023-11-15 Simulations/rbf/stencil.py

@@ -22,10 +22,8 @@ class Stencil:
             self.dim = shape[1]
         self.points = points
 
-        self.upper_bounds = np.max(points, axis=0, keepdims=True).flatten()
-        self.lower_bounds = np.min(points, axis=0, keepdims=True).flatten()
-        self.mids = (self.upper_bounds + self.lower_bounds) / 2
-        self.scalings = (self.upper_bounds - self.lower_bounds)*2
+        self.center = self.points[0]
+        self.scaling = np.max(np.abs(self.points - self.center))
 
     @property
     @cache
@@ -33,7 +31,7 @@ class Stencil:
         return len(self.points)
 
     def shift_points(self, points: np.ndarray) -> np.ndarray:
-        return (points - self.mids)/self.scalings
+        return (points - self.center)/self.scaling
 
     @property
     @cache
